---
title: "P8105_hw1_ykk2116"
author: "Yaa Asantewaa Klu"
date: "9/19/2018"
output: github_document
---

### Problem 1 

**Creating data frame 1**

```{r}
library(tidyverse)

set.seed(758)

example_df1 = tibble(
  random_sample = runif(10, min = 1, max = 5),
  vec_logical = random_sample  > 2,
  vec_char = c("Liberia", "Tanzania", "Bangladesh", "India", "Ghana", "Indonesia","Nigeria", "Togo", "China", "Thailand"),
  vec_factor = factor(c("Africa", "Africa", "Asia", "Asia", "Africa", "Asia", "Africa", "Africa", "Asia", "Asia"))
)

```


**Mean of each variable in data frame**

```{r}
mean(example_df1$random_sample)

mean(example_df1$vec_logical)

mean(example_df1$vec_char)

mean(example_df1$vec_factor)
```

_The mean of the random sample is 3.64801, whiles for the logical vector is 0.9_

_The mean of the character varaibles and factor variables could not be computed because they are not numeric or logical. Thus, do not have assigned numeric values, which is needed in the computation of a mean value. A mean can ony be computated from numbers_


**Conversion of variables from one form to the other (1)**

```{r, results='hide'}
as.numeric(example_df1$vec_logical)

as.numeric(example_df1$vec_char)

as.numeric(example_df1$vec_factor)

```

_The logical variables was converted to a numeric where TRUE was assigned a numberic value of 1 and FALSE was assigned a numeric value of 0_

_The character variables were all converted to NAs, which was introduced by coercion_

_The factor variables were all converted to 1s and 2s. 1 was assigned to Africa and 2 was assigned to Asia_


**Conversion of variables from one form to the other (2)**

```{r}
character_factor = as.factor(example_df1$vec_char)
character_factor
as.numeric(character_factor)

factor_character = as.character(example_df1$vec_factor)
factor_character
as.numeric(factor_character)

```

_The character variables were eventually converted to numeric variables, by assigning numbers 1 to 10 in the order of how the variables(countries) were arranged in the data frame. This conversion was successful because characters were first converted to factors, which can be converted to numeric variables_

_The factor variables could not be converted to numeric variables. Instead, they were forced by coercison into NAs. This is because, the factor variables were coverted into character variables, which cannot be converted into numeric variables_


### Problem 2

**Creating data frame 2**

```{r}

set.seed(433)

example_df2 = tibble(
  x = rnorm(1000),
  y = rnorm(1000),
  vec_logical = x + y > 0,
  num_vec = as.numeric(vec_logical),
  fac_vec = as.vector(vec_logical),  
)

```


**Description of vector using inline Rcode**

The size of the dataset is `r nrow(example_df2)` and the mean of x is `r mean(example_df2$x)`. The median of x is `r median (example_df2$x)` and the proportion of cases for which the logical vector is TRUE is `r (sum(example_df2$vec_logical)/nrow(example_df2))` 


**Making of scatterplots**

```{r}
scatterplot_1=ggplot(example_df2, aes(x = x, y = y, color=vec_logical)) + geom_point()

scatterplot_1
```

_The colour scales for the first plot (logical variables) is in a binary form becaus the logical variables are binary format (TRUE or FALSE)_


```{r}
scatterplot_2=ggplot(example_df2, aes(x = x, y = y, color=num_vec )) + geom_point()

scatterplot_2
```

_The colour scale for the second plot (numeric variables) are shaded or categorized in 5 different levels of intensity, from 1 to 0, with 1 being the lighest intensity and 0 being the darkest intensity_


```{r}
scatterplot_2=ggplot(example_df2, aes(x = x, y = y, color=fac_vec )) + geom_point()

scatterplot_2
```

_The colour scales for the third plot (factor variables) is in a binary form because the factor variables are in a binary format (TRUE or FALSE)_

```{r}

ggsave("first_scatter_plot.pdf", plot=scatterplot_1, height = 4, width = 6)

```



